
%%%%%%%%%%%%%%%%%%%%%%%% referenc.tex %%%%%%%%%%%%%%%%%%%%%
% sample references
% 
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% BibTeX users please use
% \bibliographystyle{}
% \bibliography{}
%
\biblstarthook{}

\begin{thebibliography}{99.}%
% and use \bibitem to create references.
%
% Use the following syntax and markup for your references if 
% the subject of your book is from the field 
% "Mathematics, Physics, Statistics, Computer Science"
%
% Contribution 
%\bibitem{science-contrib} Broy, M.: Software engineering --- from auxiliary to key technologies. In: Broy, M., Dener, E. (eds.) Software Pioneers, pp. 10-13. Springer, Heidelberg (2002)
%
% Online Document
%\bibitem{science-online} Dod, J.: Effective substances. In: The Dictionary of Substances and Their Effects. Royal Society of Chemistry (1999) Available via DIALOG. \\
%\url{http://www.rsc.org/dose/title of subordinate document. Cited 15 Jan 1999}
%
% Monograph
%\bibitem{science-mono} Geddes, K.O., Czapor, S.R., Labahn, G.: Algorithms for Computer Algebra. Kluwer, Boston (1992) 
%
% Journal article
%\bibitem{science-journal} Hamburger, C.: Quasimonotonicity, regularity and duality for nonlinear systems of partial differential equations. Ann. Mat. Pura. Appl. \textbf{169}, 321--354 (1995)
%
% Journal article by DOI
%\bibitem{science-DOI} Slifka, M.K., Whitton, J.L.: Clinical implications of dysregulated cytokine production. J. Mol. Med. (2000) doi: 10.1007/s001090000086 
%

%% general references


%%
\bibitem{anderson1991}
Anderson, W.J.: Continuous-Time Markov Chains, An Applications-Oriented Approach. Springer Series in Statistics. Springer, New York (1991)

%%
\bibitem{chap_pgm:itip}
Antonucci, A., de Campos, C.P., Zaffalon, M.: {Probabilistic graphical models}. In: Augustin, T., Coolen, F.P.A., De Cooman, G., Troffaes, M.C.M. (eds.) Introduction to Imprecise Probabilities, {John Wiley \& Sons} (2014) 

%%
\bibitem{augustin:2014}
Augustin, T., Coolen, F.P.A., De Cooman, G., Troffaes, M.C.M.: {Introduction to Imprecise Probabilities}. John Wiley \& Sons (2014) 


%%
\bibitem{benavoli2011}
Benavoli, A., Zaffalon, M., Miranda, E.: {Robust filtering through coherent lower previsions}. IEEE Transactions on Automatic Control. \textbf{56}(7), 1567--1581 (2011)


%%
\bibitem{couso:2000}
 Couso, I., Moral, S., Walley, P.: {A survey of concepts of independence for imprecise probabilities}. Risk, Decision and Policy. \textbf{5}(2), 165--181 (2000)

%%
\bibitem{cozman:2000}
Cozman, F.G.: {Credal networks}. Artificial Intelligence. \textbf{120}, 199--233 (2000)

%%
\bibitem{cozman:2005}
Cozman, F.G.: {Graphical models for imprecise probabilities}.
Int. J. Approx. Reason. \textbf{39}(2-3), 167--184 (2005)


%%
\bibitem{crossman:2009}
Crossman, R.J., Coolen-Schrijner, P., Coolen, F.P.: {Time-homogeneous birth-death processes with probability intervals and absorbing state}. J. Stat. Theory and Practice. \textbf{3}(1), 103--118 (2009)



%%
\bibitem{de2015credal}
De~Bock, J.: {Credal networks under epistemic irrelevance: theory and
  algorithms}. PhD thesis, Ghent University (2015)

%%
\bibitem{DeBock:2016}
De~Bock, J.: {The limit behaviour of imprecise continuous-time Markov chains}. J. Nonlinear Science. \textbf{27}(1), 159--196 (2017)

%%
\bibitem{debock:2017}
De Bock, J.: {Credal networks under epistemic irrelevance}. Int. J. Approx. Reason. \textbf{85}, 107--138 (2017)

%%
\bibitem{DeBock:2014ts}
De~Bock, J., De~Cooman, G.: {An efficient algorithm for estimating state sequences in imprecise hidden Markov models}. J. Artificial Intelligence Research. \textbf{50}, 189--233 (2014)

%%
\bibitem{decooman:2008}
De Cooman, G., Hermans, F.: {Imprecise probability trees: Bridging two
theories of imprecise probability}. Artificial Intelligence. \textbf{172}(11), 1400--1427 (2008)

%%
\bibitem{decooman:2009}
De Cooman, G., Hermans, F., Quaeghebeur, E.: {Imprecise Markov chains and their limit behavior}. Probability in Engineering and Informational Sciences. \textbf{23}(4), 597--635 (2009)

%%
\bibitem{deCooman:2010gd}
De Cooman, G., Hermans, F., Antonucci, A., Zaffalon, M.: {Epistemic irrelevance in credal nets: the case of imprecise Markov trees}. Int. J. Approx. Reason. \textbf{51}(9), 1029--1052 (2010)

%%
\bibitem{decooman:2015}
De Cooman, G., De Bock, J., Lopatatzidis, S.: {A pointwise ergodic theorem for imprecise Markov Chains}. In: Proceedings of ISIPTA 2015, pp. 107--115 (2015) 

%%
\bibitem{decooman:2016}
De Cooman, G., De Bock, J., Lopatatzidis, S.: {Imprecise stochastic processes in discrete time: global models, imprecise Markov chains, and ergodic theorems}. Int. J. Approx. Reason. \textbf{76}, 18--46 (2016)


%%
\bibitem{destercke:2008}
Destercke, S., De Cooman, G.: {Relating epistemic irrelevance to event trees}. 
Soft Methods for Handling Variability and Imprecision. 66--73, Springer (2008)


%%
\bibitem{erreygers2017imprecise}
Erreygers, A., De~Bock, J.: {Imprecise continuous-time Markov chains: Efficient computational methods with guaranteed error bounds}. In: Proceedings of ISIPTA 2017, pp. 145--156 (2017)

%%
\bibitem{galdino:2013}
Galdino, S.: {Interval continuous-time Markov chains simulation}. In: Proceedings of the 2013 International Conference on Fuzzy Theory and Its Applications, pp. 273-278 (2013)

%%
\bibitem{guo:2003}
Guo, X., Hern{\'a}ndez-Lerma, O.: {Continuous-time controlled Markov chains}. Ann. Appl. Probab. \textbf{13}(1), 363--388 (2003)

%%
\bibitem{guo:2009}
Guo, X., Hern{\'a}ndez-Lerma, O.: {Continuous-Time Markov Decision Processes}. Springer (2009)


\bibitem{hartfiel:1991}
Hartfiel, D.J.: {Sequential limits in Markov set-chains}.
Journal of Applied Probability. \textbf{28}(4), 910--913 (1991)

\bibitem{hartfiel:1998}
Hartfiel, D.J.: {Markov Set-Chains}. Lecture Notes in Mathematics 1695, Springer (1998)

\bibitem{hartfiel:1994}
Hartfiel, D.J., Seneta, E.: {On the theory of Markov set-chains}.
Advances in Applied Probability. \textbf{26}, 947--964 (1994)

%%
\bibitem{hermans:2012}
Hermans, F., De Cooman, G.: {Characterisation of ergodic upper transition operators}. Int. J. Approx. Reason. \textbf{53}(4), 573--583 (2012)

%%
\bibitem{augustin2013:itip}
Hermans, F., {\v{S}}kulj, D.: {Stochastic processes}.
In: Augustin, T., Coolen, F.P.A., De Cooman, G., Troffaes, M.C.M. (eds.) Introduction to Imprecise Probabilities, {John Wiley \& Sons} (2014) 


%%
\bibitem{itoh:2007}
Itoh, H., Nakamura, K.: {Partially observable Markov decision processes
with imprecise parameters}. Artificial Intelligence. \textbf{171}, 453--490 (2007)


%%
\bibitem{kozine:2002}
Kozine, I.O., Utkin, L.V.: {Interval-valued finite Markov chains}.
Reliable Computing. \textbf{8}(2), 97--113 (2002)


%%
\bibitem{krak2016ictmc}
Krak, T., De~Bock, J., Siebes, A.: {Imprecise continuous-time {M}arkov chains}. Int. J. Approx. Reason. \textbf{88}, 452--528 (2017)

%%
\bibitem{krak2017icthmc}
Krak, T., De~Bock, J., Siebes, A.: {Efficient computation of updated lower expectations for imprecise continuous-time hidden {M}arkov chains}. In: Proceedings of ISIPTA 2017, pp. 193--204 (2017)


\bibitem{stavros:phd:2017}
Lopatatzidis, S.: {Robust Modelling and Optimisation in Stochastic Processes Using Imprecise Probabilities, with an Application to Queueing Theory}.
 PhD thesis, Ghent University (2017)

%%
\bibitem{lopatatzidis:2015:isipta}
Lopatatzidis, S., De Bock, J., De Cooman, G.: {Calculating bounds on expected return and first passage times in finite-state imprecise birth-death chains}. In: Proceedings of ISIPTA 2015, pp. 177--186 (2015)

%%
\bibitem{lopatatzidis:2015b:isipta}
Lopatatzidis, S., De Bock, J., De Cooman, G.: {Computational methods for imprecise continuous-time birth-death processes: a preliminary study of flipping times}. In: Proceedings of ISIPTA 2015, p. 344 (2015)

%%
\bibitem{lopatatzidis2016queue}
Lopatatzidis, S., De~Bock, J., De~Cooman, G., De~Vuyst, S., Walraevens, J.: {Robust queueing theory: an initial study using imprecise probabilities}. Queueing Systems. \textbf{82}(1-2), 75--101 (2016)

%%
\bibitem{lopatatzidis:2017:journal}
Lopatatzidis, S., De Bock, J., De Cooman, G.: {Computing lower and upper expected first passage and return times in imprecise birth-death chains}. Int. J. Approx. Reason. \textbf{80}, 137--173 (2017) 



%%
\bibitem{chap_irr_ind:itip:2014}
Miranda, E., De Cooman, G.: {Structural judgements}. In: Augustin, T., Coolen, F.P.A., De Cooman, G., Troffaes, M.C.M. (eds.) Introduction to Imprecise Probabilities, {John Wiley \& Sons} (2014) 


%%
\bibitem{norris1998markov}
Norris, J.R.: {Markov chains}. Cambridge university press (1998)


%%
\bibitem{pearl:1988}
Pearl, J.: {Probabilistic  reasoning  in  intelligent  systems: networks  of  plausible  inference}. Morgan  Kaufmann (1988)


%%
\bibitem{rottondi:2017}
Rottondi, C., Erreygers, A., Verticale, G., De Bock, J.: {Modelling spectrum assignment in a two-service flexi-grid optical link with imprecise continuous-time Markov chains}. In: Proceedings of DRCN 2017, pp. 39--46 (2017)


%%
\bibitem{satia:1973}
Satia, J.K., Lave, R.E.: {Markovian decision processes with uncertain transition probabilities}. Operations Research. \textbf{21}, 728--740 (1973)

%%
\bibitem{skulj:2006}
{\v{S}}kulj,  D.: {Finite  discrete  time  Markov  chains  with  interval  probabilities}. In: Lawry, J, Miranda, E., Bugarin, A., Li, S., Gil, M.A.,  Grzegorzewski,  P., Hryniewicz, O. (eds.)
Soft Methods for Integrated Uncertainty Modelling, pp. 299--306, Springer (2006)

%%
\bibitem{skulj:2007}
{\v{S}}kulj, D.: {Regular finite Markov chains with interval probabilities}. In: Proceedings of ISIPTA 2007, pp. 405--413 (2007)

%%
\bibitem{Skulj:2015cq}
{\v{S}}kulj, D.: {Efficient computation of the bounds of continuous time
  imprecise Markov chains}. Appl. Math. Comp. \textbf{250}(C), 165--180 (2015)

%%
\bibitem{skulv:2013}
{\v{S}}kulj, D., Hable, R.: {Coefficients of ergodicity for Markov chains with uncertain parameters}. Metrika. \textbf{76}(1), 107--133 (2013)

%%
\bibitem{soullard2017}
Soullard, Y., Antonucci, A., Destercke, S.: {Technical Gestures Recognition by Set-Valued Hidden Markov Models with Prior Knowledge}. In: Soft Methods for Data Science, pp. 455-462 (2017)


%%
\bibitem{troffaes2015using}
Troffaes, M., Gledhill, J., {\v{S}}kulj, D., Blake, S.: {Using imprecise continuous time Markov chains for assessing the reliability of power networks with common cause failure and non-immediate repair}. In: Proceedings of ISIPTA 2015, pp. 287--294 (2015)


%%
\bibitem{vanloan:1975}
Van Loan, C.F.: {A Study of the Matrix Exponential}, Numerical Analysis Report No. 10, University of Manchester, Manchester, UK, August 1975, {Reissued as MIMS EPrint 2006.397}, Manchester Institute for Mathematical Sciences, The University of Manchester, UK.


%%
\bibitem{chap:game_theo:itip}
Vovk, V., Shafer, G.: {Game-theoretic probability}. 
In: Augustin, T., Coolen, F.P.A., De Cooman, G., Troffaes, M.C.M. (eds.) Introduction to Imprecise Probabilities, {John Wiley \& Sons} (2014) 


%%
\bibitem{Walley:1991vk}
Walley, P.: {Statistical reasoning with imprecise probabilities}. Chapman and Hall, London (1991)

%%
\bibitem{white:1994}
White, C.C., Eldeib,  H.K.: {Markov decision-processes with imprecise transition-probabilities}. Operations Research. \textbf{42}, 739--749 (1994)


\end{thebibliography}
